# Drawing Five Common Plots by ggplot2

Xirui Guo

```{r}
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(reshape2)
library(Lock5withR)
library(fivethirtyeight)
library(RColorBrewer)
library(plotly)
```

## Motivation
The ggplot2 is a powerful and effective package for drawing plots. It has a lot of syntaxes to support people get the graphs they want. A good plot needs to clearly represent the information. Scatter plots, line plots, histograms, boxplots, and heatmaps are frequently used in daily life. This Markdown hopes to be a guide about how to quickly and suitably draw the above five plots by ggplot2.

## Graph1 scatter plot: geom_point\
Using data `mtcars`\
The main code for drawing scatter plot is geom_point; however, we usually don't use it single because we need the plot shows more information and here are some syntaxes people use together with `geom_point`:\
`aes(size , color)`: sometimes it is effective, but people need to avoid showing too much information and making the graph hard to read..\
`alpha`: change the transparency of point.\
`geom_text`: add labels of each point.\
`geom_smooth`: show the regression line and correspond standard error boundary.\

```{r}
mtcars$vs <- as.factor(mtcars$vs)

ggplot(mtcars, aes(x=wt, y=mpg, size=cyl, color=vs)) +
  geom_point(alpha = .4) + 
  geom_text(label=rownames(mtcars), 
            nudge_x = 0.25, nudge_y = 0.25, 
            check_overlap = T)+
  geom_smooth(method=lm , color="red", fill="lightblue", se=TRUE)
```

## Graph2 line chart: geom_line\
People need to clarify the x-axis and y-axis and can use `group` to draw the line with different colors.
```{r}
# library(reshape2)
x <- -10:10
var1 <- dnorm(x,-2,1)
var2 <- dnorm(x,2,1)
var3 <- dt(x,2,2)
data <- data.frame(x,var1,var2,var3)
data_n <- melt(data, id="x")

ggplot(data_n,aes(x=x, y=value, group=variable, color=variable))+
  geom_line()
```

Using geom_line twice: a way for representing line from different groups without legend.
```{r}
data_n <- data_n %>%
  mutate(variable2=variable) 

ggplot(data_n, aes(x=x, y=value))+
  geom_line(data=data_n %>% select(-variable), aes(group=variable2), color="grey", size=0.5, alpha=0.8) +
  geom_line(aes(color=variable), color="blue", size=1)+
  facet_wrap(~variable)
```

## Graph3 box plot: geom_boxplot\
Using the Data: *StudentSurvey* in **Lock5withR** package(PSet1). Drawing a "simple" box-plot first:\
`labs()`: simultaneously add the main title and axis labels
```{r}
# library(Lock5withR)
p3<-ggplot(StudentSurvey, aes(x = SAT, y = Year)) + 
  geom_boxplot()+
  labs(title="Plot of SAT score and Year", x="SAT score", y="Year")
p3
```

How to make boxplot shows more information:\
1. Using `theme()` + `element_text()` to modify the color, font, size of text.
```{r}
p3+theme(
  plot.title = element_text(color = "red", size = 14, face = "bold.italic"),
  axis.title.x = element_text(color="blue", size = 14, face = "bold"),
  axis.title.y = element_text(color="green", size = 14, face = "bold")
)
```

2. Adding legend and modifying the legend tittle 
```{r}
p3<-ggplot(StudentSurvey, aes(x = SAT, y = Year, fill=Year)) + 
  geom_boxplot()+
  labs(fill="School Year")
p3
```

3.Changing the position of legend: five choices "left","top", "right", "bottom", "none"
```{r}
p3<- p3+
  theme(legend.position = "top")
p3
```

4. Modifying the title, label and background of legend
```{r}
p3+
  theme(
    legend.title = element_text(color="blue"),
    legend.text = element_text(color="red"),
    legend.background = element_rect(fill="lightblue")
    )
```

5. Changing the name and order of factor's level 
```{r}
Y <- fct_recode(StudentSurvey$Year, "NA" = "", "1st" = "FirstYear", 
                "2nd" = "Sophomore", "3rd" = "Junior", "4th" = "Senior")
Y <- fct_relevel(Y,"1st","2nd","3rd","4th")

p3<-ggplot(StudentSurvey, aes(x = SAT, y = Y, fill=Y)) + 
  geom_boxplot()+
  labs(fill="School Year")
p3
```

6. 6 ways to change the background of whole plots
```{r}
p3_1<-p3+theme_gray()
p3_2<-p3+theme_bw()
p3_3<-p3+theme_linedraw()
p3_4<-p3+theme_light()
p3_5<-p3+theme_minimal()
p3_6<-p3+theme_classic()

ggarrange(p3_1,p3_2,p3_3,p3_4,p3_5,p3_6,nrow=3,ncol=2)
```


## Graph4 histogram: geom_histogram\
Using Data: *bad_drivers* in **fivethirtyeight ** package. Common syntaxes with geom_histogram:\
`binwidth`: control the width of each bin.\
`fill`: the color of bin.\
`color`: the frame color of bin.\
`alpha`: control the transparency of the bin color.
```{r}
# library(fivethirtyeight)
p4 <- ggplot(bad_drivers,aes(x=perc_alcohol,y=..count..)) +
  geom_histogram(binwidth=3, fill="lightblue", color="black", alpha=0.7)
p4
```

`coord_flip()`：make the histogram bar become horizontal
```{r}
p4<- p4 +
  coord_flip()
p4
```

`scale_x_reverse()` and `scale_y_reverse()`：reverse the x-axis and y-axis
```{r}
p4+
  scale_y_reverse()
```

## Graph5 heatmap: geom_tile
The data is from the class notes\
```{r}
grade <- rep(c("first", "second", "third", "fourth"), 3)
subject <- rep(c("math", "reading", "gym"), each = 4)
statescore <- sample(50, 12) + 50
df <- data.frame(grade, subject, statescore)

p5 <- ggplot(df, aes(grade, subject, fill = statescore)) + 
  geom_tile() 
p5
```

1. Changing the color
`scale_fill_gradient()`, `scale_fill_distiller()` and `scale_fill_viridis()`\
`scale_fill_gradient()` can customize the color\
`scale_fill_distiller()` usually need palette = RColorBrewer\
`scale_fill_viridis()` need to let discrete=False when the variable is continuous
```{r,fig.height = 8, fig.width = 6}
# library(RColorBrewer)
display.brewer.all()
```

```{r,fig.height = 3, fig.width = 6}
p5_1<-p5 + scale_fill_gradient(low="white", high="purple") + theme(legend.position="top")
p5_2<-p5 + scale_fill_distiller(palette = "RdBu")+ theme(legend.position="top")
p5_3<-p5 + scale_fill_viridis_c()+ theme(legend.position="top")
ggarrange(p5_1,p5_2,p5_3,nrow=1,ncol=3)
```

2. adding text in each square and interact
```{r}
# library(plotly)

df <- df %>%
  mutate(text = paste("grade ", grade , "\n", "subject: ", subject, "\n", "statescore: ",statescore))
pp5<-ggplot(df, aes(grade, subject, fill = statescore, text = text)) + 
  geom_tile() 
ggplotly(pp5, tooltip="text")

```

## Effective ways for showing more information
### Way1: Faceting
1. faceting by single discrete variable：
```{r}
# vertical faceting
p3 + 
  facet_grid(Award~.)
```

```{r}
# horizontal faceting
p3 + 
  facet_grid(.~Award)
```

2. faceting by two discrete variables：
```{r}
# column facet by Year and row facet by Award
p3+
  facet_grid(Year~Award, scales='free')
```

### Way2: Representing multiple charts on a single page
Using Data: *bad_drivers* with syntaxes grid.arrange()
```{r}
# Make 3 simple graphics:
g6_1 <- ggplot(mtcars, aes(x=qsec)) + 
  geom_density(fill="slateblue")

g6_2 <- ggplot(mtcars, aes(x=drat, y=qsec, color=cyl)) + 
  geom_point(size=5) + theme(legend.position="none")

g6_3 <- ggplot(mtcars, aes(x=factor(cyl), y=qsec, fill=cyl)) + 
  geom_boxplot() +
  theme(legend.position="none")

# Plots
grid.arrange(g6_1, arrangeGrob(g6_2, g6_3, ncol=2), nrow = 2)
grid.arrange(g6_1, g6_2, g6_3, nrow = 3)
```
