# Cheatsheet for multiple graphics

Guo Pei

## Description

The cheatsheet includes a single formula sheet and a more detailed tutorial of implementing different plot types on several data sets. To be specific, it contains Histogram, Boxplot, Violin plot, Ridgeline plot, Q-Q plot, Bar chart, Cleveland dot plot, Scatterplot, Parallel coordinates plot, Biplot, Mosaic plot, Alluvial diagram and Heatmap.

For the formula sheet part, it contains nearly all formulas professor introduced in class and we used and met in the previous problem sets.

Link: https://github.com/gloria6661/5293_CC/blob/main/cheatsheet.pdf

For the implementation part, each figure is attached with code on how to draw it. For some types of plots, it lists more than one methods to draw.

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(gridExtra)
library(ggridges)
library(carData)
library(forcats)
library(dplyr)
library(tidyr)
library(tibble)
library(openintro)
library(plotly)
library(GGally)
library(scales)
library(parcoords) # devtools::install_github("timelyportfolio/parcoords")
library(d3r)
library(redav)
library(grid)
library(vcd)
library(vcdExtra)
library(ggalluvial)
```

## Histogram

Data: `iris`
```{r}
head(iris)
```

### Frequency (count) histogram (ggplot2)
```{r}
ggplot(iris, aes(Sepal.Length)) +
  geom_histogram(color = "blue", fill = "lightblue", binwidth = .3) +
  theme_grey(14) +
  labs(title = "Histogram of Sepal Length", x = "Sepal Length")
```

### Histograms with right closed / right open (base R)
```{r}
par(mfrow = c(1, 2))
# histogram with right closed
hist(iris$Sepal.Length, col = "lightblue", right = TRUE,
     breaks = 4, ylim = c(0, 60),
     main = "Histogram with right closed", xlab = "Sepal Length")
# histogram with right open
hist(iris$Sepal.Length, col = "lightblue", right = FALSE,
     breaks = 4, ylim = c(0, 60),
     main = "Histogram with right open", xlab = "Sepal Length")
```

### Density histogram with density curve overlaid (ggplot2)
```{r}
ggplot(iris, aes(x = Sepal.Length, y = ..density..)) + 
  geom_histogram(binwidth = .5, color = "blue", fill = "lightblue", boundary = 0) +
  geom_density(color = "red") +
  labs(title = "Density Histogram", x = "Sepal Length")
```

### Density histogram with density curve and theoretical normal curve overlaid (base R)
```{r}
# draw the density histogram
hist(iris$Sepal.Length, freq = FALSE, ylim = c(0, 0.5), 
     main = "Density Histogram", xlab = "Sepal Length")
# add density curve
lines(density(iris$Sepal.Length), col = 2)
# add normal curve
x <- seq(3, 9, length = 100) # x-axis grid
nc <- dnorm(x, mean = mean(iris$Sepal.Length), sd = sd(iris$Sepal.Length)) #normal curve
lines(x, nc, col = 3)
```

### Cumulative frequency histogram
```{r}
g1 <- ggplot(iris, aes(x = Sepal.Length)) +
  geom_histogram(color = "blue", fill = "lightblue") +
  labs(title = "Frequency", x = "Sepal Length")
g2 <- ggplot(iris, aes(x = Sepal.Length)) +
  geom_histogram(aes(y = cumsum(..count..)),
                 color = "blue", fill = "lightblue") +
  labs(title = "Cumulative Frequency", x = "Sepal Length")
grid.arrange(g1, g2, nrow = 1)
```

## Boxplot

### Boxplot (base R)
```{r}
boxplot(Sepal.Length ~ Species, data = iris, horizontal = TRUE,
        main = "Sepal Length vs Species", ylab = "Sepal Length")
```

### Boxplot (ggplot2)
```{r}
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  geom_boxplot(varwidth = TRUE) +
  coord_flip() +
  labs(title = "Sepal Length vs Species", y = "Sepal Length")
```

## Violin plot
```{r}
ggplot(iris, aes(x = Species,
                  y = Sepal.Length)) +
  geom_violin(adjust = 1.5) +
  coord_flip() +
  labs(title = "Violin Plot", y = "Sepal Length")
```

## Ridgeline plot
```{r}
ggplot(iris, aes(x = Sepal.Length, y = reorder(Species, Sepal.Length, median))) +
  geom_density_ridges(fill = "blue", alpha = .5, scale = .95) +
  labs(title = "Ridgeline Plot", x = "Sepal Length", y = "Species")
```

## Q-Q plot (quantile-quantile)

```{r}
qqnorm(iris$Sepal.Length)
qqline(iris$Sepal.Length, col = 2)
```

## Bar chart

Data: `TitanicSurvival`
```{r}
head(TitanicSurvival)
```

### Ordinal data (sort in logical order of the categories)
```{r}
ggplot(TitanicSurvival, aes(passengerClass)) +
  geom_bar(fill = "cornflowerblue") +
  ggtitle("Passenger Class") +
  labs(title = "Passenger Class", x = "")
```

### Nominal data (sort from highest to lowest count)
```{r}
ggplot(TitanicSurvival, aes(fct_infreq(passengerClass))) +
  geom_bar(fill = "cornflowerblue") +
  ggtitle("Passenger Class") +
  labs(title = "Passenger Class", x = "")
```

### Bar chart with facets
```{r}
ggplot(data = TitanicSurvival, aes(x = passengerClass)) +
  geom_bar() +
  facet_wrap(~survived, ncol = 1, scales = "free_y") +
  labs(title = "Bar chart faceted by Survival Status",
       x = "Passenger Class", y = "")
```

### Stacked bar chart
```{r}
ggplot(TitanicSurvival, aes(x = sex, fill = survived)) + 
    geom_bar()
```

### Grouped bar chart
```{r}
ggplot(TitanicSurvival, aes(x = sex, fill = survived)) +
  geom_bar(position = "dodge")
```

### Grouped bar chart with facets
```{r}
counts <- TitanicSurvival %>%
  group_by(sex, survived, passengerClass) %>%
  summarize(Freq = n()) %>%
  ungroup() %>%
  complete(sex, survived, passengerClass, fill = list(Freq = 0))
# draw the grouped bar chart
ggplot(counts, aes(x = sex, y = Freq, fill = survived)) +
  geom_col(position = "dodge") +
  facet_wrap(~passengerClass)
```

## Cleveland dot plot
```{r}
TitanicSurvival1 <- TitanicSurvival %>%
  rownames_to_column(var = "name")
head(TitanicSurvival1)
```

### Cleveland dot plot
```{r}
ts1 <- TitanicSurvival1 %>%
  filter(!is.na(age) & passengerClass == "1st" & survived == "yes" & sex == "female" &
           age >= 30 & age <= 40)

ggplot(ts1,aes(x = age, y = fct_reorder(name, age))) +
  geom_point(color = "blue") +
  ylab("")
```

### Cleveland dot plot with facets
```{r fig.height=10}
ts2 <- TitanicSurvival1 %>%
  filter(!is.na(age) & survived == "yes" & sex == "female" & age >= 30 & age <= 40)

ggplot(ts2, aes(x = age, y = reorder(name, age))) +
  geom_point(color = "blue") +
  facet_grid(.~reorder(passengerClass, -age, median)) +
  ylab("")
```

## Scatterplot

Data: `babies` in the openintro package
```{r}
head(babies)
```

### Scatterplot
```{r}
# draw the scatterplot
g <- ggplot(babies, aes(x = gestation, y = bwt)) +
  # adjust point size and add alpha blending
  geom_point(size = 1, alpha = .5)
g +
  # add the density contour lines
  geom_density_2d() +
  # add the linear model
  geom_smooth(method = 'lm', se = FALSE, col = 2)
```

### Interactive scatterplot
```{r}
g1 <- ggplot(iris, aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point()
ggplotly(g1)
```

### Scatterplot matrix
```{r}
plot(iris[,1:4], col = "#00660030", pch = 19)
```

## Parallel coordinates plot

Data: `state.x77`
```{r}
mystates <- data.frame(state.x77) %>%
  rownames_to_column("State") %>%
  mutate(Region = factor(state.region))
head(mystates)
```

### Static parallel coordinates plot
```{r}
ggparcoord(mystates, columns = 2:9, alphaLines = .5, 
           scale = "uniminmax", splineFactor = 10, groupColumn = 10) +
  geom_vline(xintercept = 2:8, color = "lightblue")
```

### Interactive parallel coordinates plot
```{r}
mystates %>%
  arrange(Region) %>%
  parcoords(
    rownames = FALSE,
    brushMode = "1D-axes",
    reorderable = TRUE,
    queue = TRUE,
    alpha = .5,
    color = list(
      colorBy = "Region",
      colorScale = "scaleOrdinal",
      colorScheme = "schemeCategory10"
      ),
    withD3 = TRUE,
    width = 800,
    height = 600
    )
```

## Biplot

Data: `attributes.xls`  
(http://www.econ.upf.edu/~michael/attributes.xls)
```{r}
ratings <- data.frame(country = c("Italy","Spain","Croatia","Brazil","Russia",
                                  "Germany","Turkey","Morocco","Peru","Nigeria",
                                  "France","Mexico","SouthAfrica"),
                      living_standard = c(7,7,5,5,6,8,5,4,5,2,8,2,4),
                      climate = c(8,9,6,8,2,3,8,7,6,4,4,5,4),
                      food = c(9,9,6,7,2,2,9,8,6,4,7,5,5),
                      security = c(5,5,6,3,3,8,3,2,3,2,7,2,3),
                      hospitality = c(3,2,5,2,7,7,1,1,4,3,9,3,3),
                      infrastructure = c(7,8,6,3,6,9,3,2,4,2,8,3,3))
head(ratings)
```

### Principal components analysis (PCA)
```{r}
pca <- prcomp(ratings[,2:7], scale. = TRUE)  
summary(pca)
```

### Biplot
```{r}
draw_biplot(ratings, fix_sign = FALSE)
```

### Biplot with calibrated axis and projection lines
```{r}
draw_biplot(ratings, "climate",  project = TRUE)
```

## Mosaic plot

### Mosaic plot with one variable
```{r}
counts1 <- TitanicSurvival %>%
  group_by(sex, survived) %>%
  summarize(Freq = n())
mosaic(~sex, direction = "v", counts1)
```

### Mosaic plot with two variables
```{r}
mosaic(survived ~ sex, counts1, direction = c("v", "h"))
```

### Mosaic plot with three variables
```{r}
mosaic(survived ~ passengerClass + sex, counts, direction = c("v", "v", "h"),
       rot_labels = c(0,0,0,90))
```

### Mosaic pairs plot
```{r}
pairs(table(TitanicSurvival[,c(1,2,4)]), highlighting = 2)
```

## Alluvial diagram

Data: `Yamaguchi87` in the vcdExtra package
```{r}
head(Yamaguchi87)
```

```{r}
ggplot(Yamaguchi87, aes(y = Freq, axis1 = Father, axis2 = Son)) +
  geom_flow(aes(fill = Father), width = 1/12) +
  geom_stratum(width = 1/12, fill = "grey80", color = "black") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limit = c("Father", "Son"), expand = c(.05, .05)) +
  scale_y_continuous(expand = c(.01, 0)) +
  guides(fill = FALSE)
```

## Heatmap

### Hexagonal heatmap
```{r}
ggplot(babies, aes(x = gestation, y = bwt)) +
  geom_hex()
```

### Square heatmap
```{r}
ggplot(babies, aes(x = gestation, y = bwt)) +
  geom_bin_2d()
```

### Heatmap with facets
```{r}
mydata <- Yamaguchi87 %>%
  group_by(Country, Father) %>% 
  mutate(Total = sum(Freq)) %>%
  ungroup()
ggplot(mydata, aes(x = Father, y = Son)) +
  geom_tile(aes(fill = Freq/Total), color = "white") +
  coord_fixed() +
  facet_wrap(~Country) +
  scale_fill_distiller(palette = "RdBu")
```