# Comparison between different visualization packages

Yudong Zhou and Yanbo Wang

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r}
library(ggplot2)
library(plotly)
library(tidyverse)
library(dygraphs)
library(leaflet)
library(ggfortify)
library(corrplot)
# require(devtools)
# install_github('recharts', 'taiyun')
library(recharts) # must be installed from source
```

## 0 -- Surface

Visualization is a very important part in R language. In order to facilitate users to complete visualization operations, many developers have concentrated themselves on the visualization research and developed a variety of practical packages. 

In this .html file, Yudong Zhou and Yanbo Wang will complete descriptions and summaries of several visualization packages. Hope this file may contribute a lot to other students' data-visualization learning process. 

Another purpose of this work proves to be early preparation for our final project. Yudong and Yanbo believe that these kinds of summaries and examples will help us quite a lot for the final project, in which we are planning to analyze and visualize multivariate variables and time series data.

## 1 -- ggplot2

This package, **ggplot2**, is supposed to be the most famous and the most commonly used visualization tool in R language. Up to now, we believe that students, who are enjoying going through the data visualization learning process, must have been quite familiar with the basic grammar and have completed several graphs with this useful package. Therefore, we will not go into too much detail here, instead, we will provide some typical examples of the **ggplot2** package.

In our final project, **ggplot2** is going to be our main tool to demonstrate basic values of variables and the relationships among them.

### Basic template that can be used for different types of plots: *ggplot(data = <DATA>, mapping = aes(<MAPPINGS>)) +  <GEOM_FUNCTION>()*

#### Scatterplot
```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_jitter(width = .5, size=1)
```

### Jitter Plot/Line Plot
```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point() + 
  geom_smooth(method="lm")
```


### Box Plot
```{r}
ggplot(mpg, aes(class, cty)) +
 geom_boxplot()
```

#### Pie Chart
```{r}
ggplot(mpg, aes(x = "", fill = factor(class))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) +
  coord_polar(theta = "y", start=0)
```

### Histogram
```{r}
ggplot(mpg, aes(displ)) +
  geom_histogram(aes(fill=class), bins = 10, col = "grey")
```

For more details about **ggplot2**, you can go to check the cheatsheets of it --- https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf


## 2 -- plotly
This package, **Plotly**, is pretty similar with the aforementioned **ggplot2**. However, compared to **ggplot2**, **plotly** is more able to generate interactive and high-quality plots.

In our final project, we may use **plotly** to produce interactive plots to make our outcomes much more readable and illustrative.

The next following two examples - interactive bar chart and boxplot - are the two useful plots we are going to include in our final project.

### Basic Bar Chart
```{r}
data <- ggplot2::diamonds %>% count(cut, clarity)
data %>% plot_ly(x = ~cut, y = ~n, color = ~clarity)
```

### Interactive boxplot
```{r}
plot_ly(diamonds, y = ~price) %>% add_boxplot(x = ~cut)
```


For more information about **plotly** of R, you can check the *Plotly R Open Source Graphing Library* --- https://plotly.com/r/

## 3 -- dygraphs
Like **plotly**, **dygraphs** is also able to draw out interactive plots using its dygraphs Javascripts charting library. One thing that makes this package stand out is **dygraphs** profoundly facilitates us to complete time series data visualization analysis.

In our final project, the data we are going to analyze are closely related with imports and exports. This special attribute natually makes the data become time series data. There is no doubt that **dygraphs** will be definitely helpful for us to complete time series analysis.

```{r}
dygraph(cbind(mdeaths, fdeaths)) %>%
  dySeries("mdeaths", label = "Male") %>%
  dySeries("fdeaths", label = "Female") %>%
  dyRangeSelector(height = 20)
```

## 4 -- leaflet
Once again, **leaflet** is also a practical package that can gives birth to interactive graph. But in this case, **leaflet** is specially designed for interactive maps.

This **leaflet** package seems special because of its unique property and function. We may use it to geographically analyze import and export data. 

```{r}
leaflet() %>%
  addTiles() %>%
  addMarkers(lat=40.8075, lng=-73.9626, popup="The location of Columbia University")
```

## 5 -- ggfortify
**ggfortify** extends **ggplot2** for plotting some popular R packages using a standardized approach.

Moreover, this package gives us some practical methods to specifically complete variable relationship analysis by building up models.

### Heatmap
```{r}
df <- mtcars[, c("mpg", "disp", "hp", "drat", "wt")]
df <- as.matrix(df)

# Heatmap
autoplot(scale(df))
```

### Scatterplot
```{r}
# Extract the data
df2 <- df[, c("wt", "mpg")]
colnames(df2) <- c("V1", "V2")
# Scatter plot
autoplot(df2, geom = 'point') +
  labs(x = "mpg", y = "wt")
```

### Diagnostic plots for Linear Models (LM)
```{r}
# Compute a linear model
m <- lm(Petal.Width ~ Petal.Length, data = iris)
# Create the plot
autoplot(m, which = 1:6, ncol = 2, label.size = 3)
```

### Diagnostic plots with Generalized Linear Models (GLM)
```{r}
# Compute a generalized linear model
m <- glm(Murder ~ Assault + UrbanPop + Rape,
         family = gaussian, data = USArrests)
# Create the plot
# Change the theme and colour
autoplot(m, which = 1:6, ncol = 2, label.size = 3,
         colour = "steelblue") + theme_bw()
```

### Plotting time series
```{r}
autoplot(AirPassengers)
```

## 6 -- corrplot
R package **corrplot** provides a visual exploratory tool on correlation matrix that supports automatic variable reordering to help detect hidden patterns among variables.
```{r}
M = cor(mtcars)
corrplot(M, method = 'number') # colorful number
```

### Change the order by 'alphabet'
```{r}
corrplot(M, method = 'color', order = 'alphabet')
```

### Change the method by 'shade'
```{r}
corrplot(M, method = 'shade', order = 'alphabet')
```

## 7 -- recharts
The R package **recharts** provides an interface to the JavaScript library ECharts for data visualization. The goal of this package is to make it easy to create charts with only a few lines of R code.

```{r}
iris$Species <- as.character(iris$Species)
ePoints(iris, ~Sepal.Length, ~Sepal.Width, series = ~Species)
```

## 8 -- Summary
In this file, considering the nature of the final project data, Yudong Zhou and Yanbo Wang have mainly completed the enumeration and simple examples of the visualization tools that may be involved. We will use them fully and reasonably in the final project to complete import and export data relationship analysis and visualization.




