# Twitter API guide in R

Jiayi Yuan

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are the libraries needed for this chapter.
```{r}
###CRAN installation
library(twitteR)

###This is a GitHub installation
#install.packages(c("devtools", "rjson", "bit64", "httr")) 
library(devtools)
#install_github("geoffjentry/twitteR")

###CRAN installation
library(rtweet)
```

## Background
API, known as application programming interface, is always a strong tool for data scientists or statisticians to gain useful data or information from different websites, computer programs, etc. I tried to collect data using Twitter API for my final project, and I noticed that even though the tool is helpful, not so many of my classmates are familiar with it or used it before, and many of the guides online are outdated. Therefore, in this community contribution, I would like to provide a comprehensive guide on using Twitter API to search tweets or gather some other information, such as followers or favorited tweets, in R. Using the Twitter API should be an easy thing, but there could still be many problems for those who are totally new to such a tool. Therefore, I would like to introduce different setup methods to make sure that everyone who reads this guide can successfully link the interface to R. I will also include some examples of R codes and the output data in this guide to show more directly how the interface works.

As a beginner, it took me a long time to set up the interface. I got many problems and tried to search the solutions online, but I found that not many people had those problems. Even though, for now, I still have many errors that are hard to find an answer online. Therefore, I believe an up-to-date guide on using Twitter API is necessary, and I would also like to introduce such a tool to more people who would like to gather data from webs and would not be freaked out by potential problems at the very beginning stage. 

Although I have not tried other APIs, I think the underlying logic is the same. By learning Twitter API, I believe I can also take advantage of getting familiar with other user APIs quickly. Therefore, learning how to use an API can be really helpful to me in my future study and I will also try more APIs or developer tools provided by other platforms

## Setup
### Setup Twitter Developer Account and APP
(1). Create a Twitter account if you don't have one, and associate your phone number to the Twitter account.

(2). Apply for a Twitter Developer Account (https://developer.twitter.com/en) with your personal Twitter account. Fill out all the information needed and follow the guidance, there shouldn't be a problem.

(3). Create a Twitter API App. This step provide you with the important credentials that will be used in the setup of following connections between R and the interface. You will need to provide an App name and write an Application description. These two sections are easy to fill out, just like creating a file on your computer, but some of the following sections maybe unfamiliar. For the Website URL, you can put the link of your personal Twitter account and this is the most recommended URL link. Make sure both the OAuth 2.0 and OAuth 1.0a are allowed in the authentication settings, or there might problems in the setup of R. You will also need to provide a Callback URL. http://127.0.0.1:1410, this link can be used as the Callback URL and it should be exactly the same link. The Callback URL makes sure you can link your R studio to your App. Then you can successfully create a Twitter API App. Once you created, you can get the keys and tokens for this App. They are important and private, make sure to copy them and do not tell others. If you lose them, you can also find them afterwards in the developer portal and you can also regenerate the keys. The keys and tokens we need for the setup in R are the consumer(API) key, consumer(API) secret, access token and access secret. 

### Setup in R
After getting the information we need from the developer account, we can now move to R.

There are two ways to link the API into Rstudio, and I am going to start with the most popular one.

**Method 1: twitterR Package**

twitterR is the package that provides access to twitter API. You can install the package either with a CRAN version or with a github version. Personally, I recommend using the CRAN version directly as the github version sometimes create unexpected errors. (The two versions of the installation code are at the begging of this chapter.)

After installing the package, the next thing is to link the API into R using the authentication keys. Using the following code to set up the authentication. While setting up the authentication, R will ask whether you would like to use a local file to cache the access credentials. If yes, enter 1 in the console and if not, enter 2. For here, I entered 2. 
After this selection, you have successfully linked your R studio to the API, and you can you R codes to get tweets now!
```{r}
####Tokens get from your Twitter Developer account
consumer_key <- "By1wB2FM6kb3oDCmLt1SKCdl5"
consumer_secret <- "tK4hKi7JxGAmOHsqVOWaGsnyRWw7JcqQjARrHu0juPIPDJOXq3"
access_token <- "1500916491105021960-hG1qpuiFQww8Gd3Q7mVtzoCSvLKTZ9"
access_secret <- "8Du6PQ0b6JRvoDHGK2x893ZUgJ4tf5d90fnZjFOJ9d5CV"
setup_twitter_oauth(consumer_key, consumer_secret, access_token, access_secret)
```

**Method 2: rtweet Package**

The second method to build to link is by using another package called rtweet. There is no big difference between these two packages, but the rtweet package is the method I found when I have troubles setting authentication with the twitteR package. Personally, I suggest using rtweet, as it is neater, more user-friendly and have more built-in functions.
The process is basically the same: installing the rtweet package, call the package and setup the authentication. For this package, we are only going to use the CRAN version. (The installation code is included at the beginning of the chapter.)

The authentication step is almost the same except the code is a little bit different. When running the following code, it will not show the prompt of asking whether to cache the access as the above code does, but there will be a pop-up window saying the authentication is successfully built.
```{r}
token <- create_token(
  app = "5293Data", 
  #app here is the name of the Twitter API App you created, and my app name is "5293Data".
  consumer_key = consumer_key,
  consumer_secret = consumer_secret,
  access_token = access_token,
  access_secret = access_secret)
```

This method automatically save your token as an environment variable in your computer. So next time you start an R session (on the same machine), rtweet should automatically find your token. You can use the following code to make sure your token is correct.
```{r}
get_token()
```

## Examples
Now we can see some sample codes of how to get Twitter information exactly.

### Get tweets
Following is an example of getting the tweets including the key words *spotify* and *down*, and convert the twitterR list to a dataframe. One thing needs to be noticed here is that although there are arguments of *since* and *until*, which allow you to search tweets in a certain range of dates, Twitter API can only provide a sample of tweets published in the past 7 days. Therefore, if you want to search the tweets 7 days ago, no result will be shown.
You could also use a function called *search_30day()* to retrieve tweets in the past 30 days, but this can only be accessed by premium APIs.
```{r}
##This is a function from package twitteR
spotify <- searchTwitter('spotify + down', n = 100, since = "2022-03-21", 
                         until = "2022-03-28")
spotify_df  <- twListToDF(spotify)
head(spotify_df)

##This is a function from package rtweet
spotify30 <- search_30day("spotify", n = 300, env_name = "5293Data", 
                          toDate = 202203080000)
```

### Get favorited tweets
This example tells you how to get favorited tweets by certain users. With the following code, we can get the most recent 300 favorited tweets by *Columbia University* and *Netflix*.
```{r}
##This is a function from package rtweet
favs <- get_favorites(c("Columbia", "Netflix"), n=300)
head(favs)
```

### Get followers
Like the title suggests, this example code provides the information of the followers of (a) certain user(s). The drawback of this function is that it only returns the user id but no other information.
```{r}
## This is a function from package rtweet
spotify_flw <- get_followers("Columbia", n = 750)
head(spotify_flw)
```

### Get information of a certain user
This sample code allows you to retrieve the base information of certain Twitter users, such as name, account created time, number of followers, lists of friends, etc. The following code shows the information of *Columbia University*, *Netflix*, *Spotify*, *Yale University*, *University of Rochester* and *Biden.*
```{r}
## This is a function from package twitteR
tuser <- getUser('Columbia')
users <- lookupUsers(c('Columbia',"Netflix","Spotify","Yale","UofR","POTUS"))
user_df <- twListToDF(users) 
user_df
```

### Visualize tweets
There is a useful function in the rtweet package allow us to visualize the frequency of tweets. The following code shows the tweet frequency of the three universities: *Columbia University*, *University of Rochester* and *Yale University* of their 300 most recent tweets.
```{r}
library(dplyr)
## This is a function from package rtweet
university <- get_timelines(user = c("Columbia", "UofR","Yale"), n=300)
university %>%
  group_by(screen_name) %>%
  ts_plot(by = "week") + 
  labs(title = "Frequency of Universities Twitter Statuses")
```

## Sources
https://cran.r-project.org/web/packages/rtweet/rtweet.pdf

https://www.rdocumentation.org/packages/twitteR/versions/1.1.9

https://medium.com/@GalarnykMichael/access-data-from-twitter-api-using-r-and-or-python-b8ac342d3efe







