---
title: "Interactive Presentation with xaringan and dygraphs"
author: "Ziwen Ye"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: center, middle

# Quick start for dygraphs
---

```{r, fig.align='center', message = FALSE}
library(dygraphs)
lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths)

```

---
class: center, middle
# Modified graphs with different opition
- Add options with dyOptions

-- colors: add color for each line

-- stepPlot: make line as step

-- drawPoints: draw point in the line



---
Add colors

```{r, fig.align='center', message = FALSE}

lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dyOptions(colors = RColorBrewer::brewer.pal(3, "Set2"))

```



---
Add stepPlot

```{r, fig.align='center', message = FALSE}

lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dyOptions(stepPlot = TRUE)

```
---


Add fillGraph

```{r, fig.align='center', message = FALSE}

lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dyOptions(fillGraph = TRUE, fillAlpha = 0.5)

```
---
Add Point

```{r, fig.align='center', message = FALSE}

lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dyOptions(drawPoints = TRUE, pointSize = 2)

```

---
class: center, middle
# Modified graph with Legends & Labels

-- main: title

-- xlab: label for x axis

-- ylab: label for y axis

---
```{r, fig.align='center', message = FALSE}

lungDeaths <- cbind(mdeaths, fdeaths)
dygraph(lungDeaths, 
        main = "Deaths from Lung Disease (UK)", 
        ylab = "Death / Year",
        xlab = "Year")

```


---

class: center, middle
# Per-Series method
---
Draw series line by line
```{r, fig.align='center', message = FALSE}

dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dySeries("fdeaths", color = "red")
```

----
Add Highlight for series
```{r, fig.align='center', message = FALSE}

dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dySeries("fdeaths", color = "red") %>%
  dyHighlight(highlightCircleSize = 5, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE)
```
---
Add two axis for series

```{r, fig.align='center', message = FALSE}

dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dyAxis("y", label = "mdeath") %>%
  dyAxis("y2", label = "fdeaths", independentTicks = TRUE) %>%
  dySeries("fdeaths", color = "red", axis = 'y2')
```

---
class: center, middle
# Annotation & Shading
---
class: center, middle
# Annotation
---
Add Annotation

```{r, fig.align='center', message = FALSE}
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dyAxis("y", label = "mdeath")  %>%
  dyAnnotation("1976-02-01", text = "Highest", width = 60)

```

---
class: center, middle
# Shading

---
Add shade for region
```{r, fig.align='center', message = FALSE}
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dyAxis("y", label = "mdeath")  %>%
  dyAnnotation("1976-02-01", text = "Highest", width = 60) %>%
  dyShading(from = "1975-11-1", to = "1976-8-1", color = "#FFE6E6")

```

---
class: center, middle
# Events & Limits
## Similar to vline and hline

---
Events
```{r, fig.align='center', message = FALSE}
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dyAxis("y", label = "mdeath") %>%
  dyShading(from = "1975-11-1", to = "1976-8-1", color = "#FFE6E6") %>%
  dyEvent("1975-11-1", "Event Start", labelLoc = "bottom") %>%
  dyEvent("1976-8-1", "Event End", labelLoc = "bottom")
```

---
Limits
```{r, fig.align='center', message = FALSE}
dygraph(lungDeaths, main = "Deaths from Lung Disease (UK)") %>%
  dySeries("mdeaths", color = "blue") %>%
  dyAxis("y", label = "mdeath") %>%
  dyShading(from = "1975-11-1", to = "1976-8-1", color = "#FFE6E6") %>%
  dyEvent("1975-11-1", "Event Start", labelLoc = "bottom") %>%
  dyEvent("1976-8-1", "Event End", labelLoc = "bottom") %>%
  dyLimit(as.numeric(lungDeaths[26,1]), color = "red", "Highest death")
```

---

class: center, middle

# Thanks!
